Methods to prioritise pop-up active transport infrastructure

### 2020-05-09 article.Rmd
Update article

<!-- Currently the intro is 500 words, too long! -->

<!-- The Covid-19 pandemic has transformed all sectors of the economy, not least transport. -->
<!-- [Demand for long distance trips has plummetted](https://osf.io/preprints/socarxiv/v3g5d/) and [airline companies have gone bust](https://arxiv.org/abs/2004.08460) [@iacus_estimating_2020; @jittrapirom_exploratory_2020]. -->
Much attention has focused on the impacts of Covid-19 on long distance travel patterns [e.g. @iacus_estimating_2020; @jittrapirom_exploratory_2020].
Yet short distance travel patterns have also changed, with a notable uptick in active travel, particularly cycling, in some areas  [@harrabin_boom_2020].
The two main explanations for this are 1) the need increased need for exercise close to home during lockdowns for mental and physical health [@jimenez-pavon_physical_2020]; and 2) reduced public transport options and use [e.g. @tian_investigation_2020].

@@

The second reason is particularly important given that many 'key workers' are low paid, with limited access to private automobiles.
From a physical activity perspective this change is welcome: obesity is a powerful predictor of all-cause mortality, including from Covid-19 [@docherty_features_2020].
From a physical distancing perspective, increased levels of walking and cycling,
the shift creates pressure on governments to ensure sufficient 'space for social distancing', while enabling increased active mobility for health and travel to key workplaces.
<!-- particularly in densly populated urban areas where pavements and cycleways may be narrow,  -->

Local and national governments worldwide are still working out how best to respond to these changes and many options are available to ensure that citizens can benefit from outdoor activity while minimizing health risks, ranging from the provision of hand sanitisers to opening-up spaces such as parking lots and golf courses [@freeman_covid19_2020].
Many local governments have responded by implementing 'pop-up' active transport infrastructure.
The Scottish government, for example has provided emergency funding "to keep key workers moving" by "reallocating road space to better enable this shift and make it safer for people who choose to walk, cycle or wheel for essential trips or for exercise" [@transportscotland_10_2020].
A wide range of interventions is possible, ranging from simple signing to the construction of new paths, but the majority of pop-up activity can be classified into three broad categories:

@@

Datasets from the PCT and CyIPT project were combined into a single file, key variables from which are shown in Table 1.
A map showing the spatial distribution of hospitals in the case study city of Leeds, which is used for demonstrating the methods in the next section, is shown in Figure \@ref(fig:hospitals).

>>>>

<!-- Currently the intro is 350 words, just about OK! -->

<!-- The Covid-19 pandemic has transformed all sectors of the economy, not least transport. -->
<!-- [Demand for long distance trips has plummetted](https://osf.io/preprints/socarxiv/v3g5d/) and [airline companies have gone bust](https://arxiv.org/abs/2004.08460) [@iacus_estimating_2020; @jittrapirom_exploratory_2020]. -->
Much attention has focused on the impacts of Covid-19 on long distance travel patterns [e.g. @iacus_estimating_2020; @jittrapirom_exploratory_2020].
Yet short distance travel patterns have also changed, with a notable uptick in active travel, particularly cycling, in some areas  [@harrabin_boom_2020].
The two main explanations for this are 1) the need increased need for exercise close to home during lockdowns for mental and physical health [@jimenez-pavon_physical_2020]; and 2) reduced public transport options and use [e.g. @tian_investigation_2020].
The second reason is particularly important given that many 'key workers' are low paid, with limited access to private automobiles.

@@

<!-- From a physical activity perspective this change is welcome: obesity is a powerful predictor of all-cause mortality, including from Covid-19 [@docherty_features_2020]. -->
<!-- From a physical distancing perspective, increased levels of walking and cycling, -->
<!-- the shift creates pressure on governments to ensure sufficient 'space for social distancing', while enabling increased active mobility for health and travel to key workplaces. -->
<!-- particularly in densly populated urban areas where pavements and cycleways may be narrow,  -->

Local and national governments are working out how best to respond.
Many options are available to ensure that citizens can benefit from outdoor activity while minimizing health risks, ranging from the hand sanitiser provision to the creation of extra active transport space [@freeman_covid19_2020].
Installation of 'pop-up' active transport infrastructure has been endorsed and implemented in many places [@laker_world_2020].
The Scottish government, for example, has provided £10m "to keep key workers moving" by "reallocating road space to better enable this shift and make it safer for people who choose to walk, cycle or wheel for essential trips or for exercise" [@transportscotland_10_2020].
Most pop-up active transport infrastructure can be classified into three broad categories:

@@

Datasets from the PCT and CyIPT project were merged, resulting key variables summarised in Table 1.
A map showing the spatial distribution of hospitals in the case study city of Leeds, which is used for demonstrating the methods, is shown in Figure \@ref(fig:hospitals).
